---
title: 'Part 2: Basic Inferential Data Analysis Instructions'
author: "Chuwei Zhong"
date: "October 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data && simple analysis
```{r eval=FALSE} 
library(ggplot2)
```

```{r ToothGrowth} 
data(ToothGrowth)

str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
dim(ToothGrowth)
table(ToothGrowth$supp,ToothGrowth$dose)

```

## basic summary

```{r} 
library(ggplot2)
ggplot(ToothGrowth, aes(factor(supp), len, fill = factor(supp))) +
      geom_boxplot() + facet_grid(.~dose) +xlab("Supplement type") + ylab("Tooth  Lengh")

```


## Use confidence intervals and/or hypothesis tests to compare tooth growth by supply and dose

Here,null hypothesis is: tooth growth has nothing to do with the supply in different levels of dosages.  


```{r} 
#compare the influence by delivery method for the same dosage 
data05 <- rbind(ToothGrowth[(ToothGrowth$dose == 0.5) & (ToothGrowth$supp == "OJ"),], ToothGrowth[(ToothGrowth$dose == 0.5) & (ToothGrowth$supp == "VC"),])
t05 <- t.test(len ~ supp, data05, var.equal = FALSE, conf.level = 0.95)

data1 <- rbind(ToothGrowth[(ToothGrowth$dose == 1) & (ToothGrowth$supp == "OJ"),], ToothGrowth[(ToothGrowth$dose == 1) & (ToothGrowth$supp == "VC"),])
t1 <- t.test(len ~ supp, data1, var.equal = FALSE, conf.level = 0.95)

data2 <- rbind(ToothGrowth[(ToothGrowth$dose == 2) & (ToothGrowth$supp == "OJ"),], ToothGrowth[(ToothGrowth$dose == 2) & (ToothGrowth$supp == "VC"),])
t2 <- t.test(len ~ supp, data2, var.equal = FALSE, conf.level = 0.95)

#extract p-values and confidence intervals
summary_result <- data.frame(
  c(t05$p.value, t1$p.value, t2$p.value), 
  c(t05$conf.int[1],t1$conf.int[1], t2$conf.int[1]),
  c(t05$conf.int[2],t1$conf.int[2], t2$conf.int[2])
)
row.names(summary_result) <- c("Dosage 0.5","Dosage 1.0","Dosage 2.0")
colnames(summary_result) <- c("p.value","Conf.Low","Conf.High")
summary_result

```

Here, for dosages to be 0.5mg and 1.0mg, p-values are less than 0.05 and confidence intervals don’t include 0. So we reject the null hypothesis. For dosage to be 2.0mg, p-values are larger than 0.05 and confidence intervals include 0. Hence, we fail to reject the null hypothesis. 

## State my conclusions and assumptions

In lower dosages(0.5-1.0mg), orange juice can improve tooth growth than Vitamin C. While in higher dosage(2.0mg), tooth growth rate is not statistically different between these two supplyment methods. 
