---
title: "Statistical Inference Peer Assignment 1"
author: "Chuwei Zhong"
date: "October 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is the report for Statistical Inference Peer Assignment 1. Also this is my first R Markdown document. 

## Simulation

```{r}
#Create an empty vector
mns = NULL
#fill the vector with 1000 exmaples
for (i in 1 : 1000) mns = c(mns, mean(rexp(40, rate=0.2)))
```

## Sample Mean versus Theoretical Mean

```{r}
#hisgram to show
hist(mns)
#sample mean versus theoretical mean
sample_mean <- mean(mns)
theoretical_mean <- 1/0.2
#putm mean value in the hisgram
abline(v=sample_mean,col="blue",lwd=3)
abline(v=theoretical_mean,col="red",lwd=3)
```
You will be able to see that two line are almost vertical and overlape with each other. 

## Sample Variance versus Theoretical Variance

```{r}
#now calc sample variance versus theoretical variance
sample_variance <- var(mns)
sample_variance
theoretical_variance <- (1/0.2^2)/40
theoretical_variance
```

## Distribution
Let's see how the distribution is approximately normal

```{r}
qqnorm(mns)
qqline(mns)
```

You can see from the plot that the distribution is quite normal. 
